Inequality in socioemotional skills: a cross-cohort comparison

### 2018-10-02 codes/R/paper/text/paper_jpube_1introlit.tex
Sent for IFS WP

There are only few studies in economics specifically studying ``non-cognitive'' traits and health behaviors. \citet{Conti2010a} and \citet{Conti2011} were the first to consider three early endowments, including child socio-emotional traits and health in addition to cognition, using rich data from the 1970 British cohort study. They find strong evidence that non-cognitive traits promote health outcomes and healthy behaviors, and than not accounting for them overestimates the effects of cognition; additionally, they document that child cognitive traits are more important predictors of employment and wages than socio-emotional traits or early health. \citet{Chiteji2010} used the US Panel Study of Income Dynamics (PSID) and found that future orientation and self-efficacy (related to emotional stability) are associated with less alcohol consumption and more exercise. \citet{Cobb-Clark2014} used the Australian HILDA data and found that an internal locus of control (also related to emotional stability, perceived control over one's life) is related to better health behaviours (diet, exercise, alcohol consumption and smoking). \citet{Mendolia2014a} used the Longitudinal Study of Young People in England and found that individuals with external locus of control, low self-esteem, and low levels of work ethics, are more likely to engage in risky health behaviours. \citet{Savelyev2017} show that the association between personality traits and health behaviours also holds in a high-IQ sample (the Terman Sample). \citet{Heckman2016a} use, instead, early risky behaviours to measure socio-emotional traits, and confirm their predictive power for health behaviours and health outcomes.

>>>>

There are only few studies in economics specifically studying ``non-cognitive'' traits and health behaviors. \citet{Conti2010a} and \citet{Conti2011} were the first to consider three early endowments, including child socio-emotional traits and health in addition to cognition, using rich data from the 1970 British cohort study. They find strong evidence that non-cognitive traits promote health outcomes and healthy behaviors, and than not accounting for them overestimates the effects of cognition; additionally, they document that child cognitive traits are more important predictors of employment and wages than socio-emotional traits or early health. \citet{Chiteji2010} used the US Panel Study of Income Dynamics (PSID) and found that future orientation and self-efficacy (related to emotional stability) are associated with less alcohol consumption and more exercise. \citet{Cobb-Clark2014} used the Australian HILDA data and found that an internal locus of control (also related to emotional stability, perceived control over one's life) is related to better health behaviours (diet, exercise, alcohol consumption anwhad smoking). \citet{Mendolia2014a} used the Longitudinal Study of Young People in England and found that individuals with external locus of control, low self-esteem, and low levels of work ethics, are more likely to engage in risky health behaviours. \citet{Savelyev2017} show that the association between personality traits and health behaviours also holds in a high-IQ sample (the Terman Sample). \citet{Heckman2016a} use, instead, early risky behaviours to measure socio-emotional traits, and confirm their predictive power for health behaviours and health outcomes.
